---
title: "Edx Movie Lens, Harvard 125.9"
author: "Doug Northcott"
date: "6/13/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Movie Lens

This project is to analyze this large data sets that is beyond the limit of most machines and subject this dataset to machine learning to find the most accurate predicting model.

```{r, echo=FALSE}


if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- read.table(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                      col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))

movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data

set.seed(1) # if using R 3.6.0: set.seed(1, sample.kind = "Rounding")
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set

validation <- temp %>% 
     semi_join(edx, by = "movieId") %>%
     semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set

removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

## Basic Stats
# The Data Set is the 10M data set from the movielens data base, prepared for a class through Harvard University Online


```{r}
summary(edx)
library(psych)
library(magrittr)
library(tidyverse)
describe(edx)
edxg <- edx %>% separate_rows(genres, sep = "\\|")
edxg %>%
	group_by(genres) %>%
	summarize(count = n()) %>%
	arrange(desc(count))
```
After the genres category has been split into individual categories, a numbers is then assigned each category to assist with the regression
```{r}
edxg[edxg$genres == "Drama", "genres"] <- 1
edxg[edxg$genres == "Comedy", "genres"] <- 2
edxg[edxg$genres == "Action", "genres"] <- 3
edxg[edxg$genres == "Thriller", "genres"] <- 4
edxg[edxg$genres == "Adventure", "genres"] <- 5
edxg[edxg$genres == "Romance", "genres"] <- 6
edxg[edxg$genres == "Sci-Fi", "genres"] <- 7
edxg[edxg$genres == "Crime", "genres"] <- 8
edxg[edxg$genres == "Fantasy", "genres"] <- 9
edxg[edxg$genres == "Children", "genres"] <- 10
edxg[edxg$genres == "Horror", "genres"] <- 11
edxg[edxg$genres == "Mystery", "genres"] <- 12
edxg[edxg$genres == "War", "genres"] <- 13
edxg[edxg$genres == "Animation", "genres"] <- 14
edxg[edxg$genres == "Musical", "genres"] <- 15
edxg[edxg$genres == "Western", "genres"] <- 16
edxg[edxg$genres == "Film-Noir", "genres"] <- 17
edxg[edxg$genres == "Documentary", "genres"] <- 18
edxg[edxg$genres == "IMAX", "genres"] <- 19
edxg[edxg$genres == "(no genres listed)", "genres"] <- 20
head(edxg)
edxg$genres <- as.double(edxg$genres)
write.csv(edxg, "C:\\Users\\dougl\\OneDrive\\Email attachments\\Documents\\edxg.csv", row.names = FALSE)
```
In preparation  for transfer into the Spark the data is backed up on Microsoft Cloud, then the Spark connection is established

```{r}
library(sparklyr)
library(dplyr)
spark_install()
sc <- spark_connect(master = "local")
```
Once the Spark connection is established then the Data is loaded into Spark

```{r}
edxgspark <- spark_read_csv(sc, name = "edxg", path = "C:\\Users\\dougl\\OneDrive\\Email attachments\\Documents\\edxg.csv")
count(edxgspark)
```
In spark you can count the number of rows which is now in excess of 20,000,000 

## The comparison of the various Regression application in the MLib library
```{r}
modellr <- edxgspark %>% ml_linear_regression(rating ~ genres + userId)
summary(modellr)
```


```{r}
modelglm <- edxgspark %>% ml_generalized_linear_regression(rating ~ genres + userId)
summary(modelglm)
```

```{r}
modelrf <- edxgspark %>% ml_random_forest(rating ~ genres, type = "regression")
summary(modelrf)
print(modelrf)
```

```{r}
modelgb <- edxgspark %>% ml_gradient_boosted_trees(rating ~ genres, type = "regression")
summary(modelgb)
```

## Preparation of the Validation Set
```{r}
vg <- validation %>% separate_rows(genres, sep = "\\|")
vg[vg$genres == "Drama", "genres"] <- 1
vg[vg$genres == "Comedy", "genres"] <- 2
vg[vg$genres == "Action", "genres"] <- 3
vg[vg$genres == "Thriller", "genres"] <- 4
vg[vg$genres == "Adventure", "genres"] <- 5
vg[vg$genres == "Romance", "genres"] <- 6
vg[vg$genres == "Sci-Fi", "genres"] <- 7
vg[vg$genres == "Crime", "genres"] <- 8
vg[vg$genres == "Fantasy", "genres"] <- 9
vg[vg$genres == "Children", "genres"] <- 10
vg[edxg$genres == "Horror", "genres"] <- 11
vg[vg$genres == "Mystery", "genres"] <- 12
vg[vg$genres == "War", "genres"] <- 13
vg[vg$genres == "Animation", "genres"] <- 14
vg[vg$genres == "Musical", "genres"] <- 15
vg[vg$genres == "Western", "genres"] <- 16
vg[vg$genres == "Film-Noir", "genres"] <- 17
vg[vg$genres == "Documentary", "genres"] <- 18
vg[vg$genres == "IMAX", "genres"] <- 19
vg[vg$genres == "(no genres listed)", "genres"] <- 20
vg$genres <- as.double(vg$genres)
write.csv(vg, "C:\\Users\\dougl\\OneDrive\\Email attachments\\Documents\\vg.csv", row.names = FALSE)
vgspark <- spark_read_csv(sc, name = "vg", path = "C:\\Users\\dougl\\OneDrive\\Email
```
## Comparison of the Random Forest and Gradient Boosted Model on the Validation Set
```{r}
predrf <- ml_predict(modelrf,vgspark)
ml_regression_evaluator(predrf, label_col = "rating")
```


```{r}
predgb <- ml_predict(modelgb,vgspark)
ml_regression_evaluator(predgb, label_col = "rating")
```

